---
title: "Mouse-tracking-analysis"
output: html_document
date: "2022-08-12"
author: "Swarnendu Moitra, Indranil Dutta"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear environment
rm(list = ls())
```

```{r}
# Load package
library("mousetrap")
```

# CREATE MOUSE TRACKING OBJECT FROM DATA FILE

## OPEN DATA FILE AND SAVE IT AS DATAFRAME OBJECT "d" (select the "TEST_MT.csv" file in the file browser pop-up)
  # The file must contain 3 columns that correspond to the x and y successive positions of the mouse and corresponding
  # time stamps. In each cell, the values are separated by a comma. The separator for column must be different (i.e. semi-colons)
  # For any given row (i.e. trial), those three cells must contain a list of values of equal length.

```{r}
#d=read.csv(file.choose(),sep=";", header = T)

d=read.csv(file.choose(),sep=";", header = T)
e=read.csv(file.choose(), header = T)
```

## Create mouse tracking object from data frame
   indicate the data frame, and which columns in it correspond to the x positions, the y positions,
   the time stamps, and the column that serves as identifier for the trial)

### Observe the mouse tracking object
    NA values appear because different trials are not of the same duration. This is solved by time normalization later.
```{r}
library(dplyr)
d <- d %>% rename(Trial = Trial)
d <- filter(d, Parameter == "Move")
```


# Trial = Trial
```{r}
mtdata = mt_import_mousetrap(
  d,
  xpos_label = "xpos",
  ypos_label = "ypos",
  timestamps_label = "timestamps",
  mt_id_label = "Trial",
)

View(mtdata)
```

## Plot the trajectories
   the TEST_MT.csv file provides data for 6 trials, so 6 trajectories appear
   note that during the experiment, participants' mouse start at the bottom of the screen and move up to one of two
   answer options situated in the upper corners of the screen. The software used to run the experiment places y = 0
   at the top of the screen, so the trajectories are inverted vertically. This is solved by axis reorganization later.

```{r}
mt_plot(mtdata)
```

This line can group by color different trajectories with an identical value on the column indicated in the "color" argument (can be ignored)

```{r}
mt_plot(mtdata, use = "trajectories", use2 = "data", color="Trial")
```

# ALIGN AND NORMALIZE TRAJECTORIES 
To be comparable, each trajectories must have identical starting and ending position and be normalized in space
and time. The following code makes all the necessary adjustments. Plotting after every steps allows to visualize
the adjustments

## Align the starting position of each trajectories
```{r}
mtdata <- mt_align_start(
  mtdata,
  use = "trajectories",
  save_as = "trajectories",
  dimensions = c("xpos", "ypos"),
  start = c(0,0),
  verbose = FALSE
)
```


```{r}
# Plot the trajectories (start position aligned)
mt_plot(mtdata)

# Plot with colors (can be ignored)
mt_plot(mtdata, use = "trajectories", use2 = "data", color="Trial")
```

# Space normalize (align start and end positions)

```{r}
mtdata = mt_align_start_end(
  mtdata,
  use="trajectories",
  save_as = "sn_trajectories",
  dimensions=c("xpos","ypos"),
  start=c(0,0),
  end=c(-1,1),
  verbose=FALSE
)
```

# Plot the trajectories (with aligned start/end position)
  # Note that all trajectories now go in the same direction

```{r}
mt_plot(mtdata$sn_trajectories)
```

# Plot with colors (can be ignored)

```{r}
mt_plot(mtdata, use = "sn_trajectories", use2 = "data", color = "Trial")
```

# Time normalize
  # All trajectories are normalized to be composed of 101 (a default value) time steps of equal size,
  # thus normalizing trials of different duration in terms of time
  
```{r}
mtdata = mt_time_normalize(
  mtdata,
  use="sn_trajectories",
  save_as = "tn_trajectories",
  dimensions=c("xpos","ypos"),
  timestamps="timestamps",
  nsteps=101,
  verbose=FALSE
)
```



# Reorganize axis (puts the origin at the bottom left)

```{r}
mtdata = mt_remap_symmetric(
  mtdata,
  use = "tn_trajectories",
  save_as = "tn_trajectories",
  dimensions = c("xpos", "ypos"),
  remap_xpos = "right",
  remap_ypos = "up"
)
```




# Plot trajectories (time normalized)

```{r}
mt_plot(mtdata$tn_trajectories)
```


# Plot with colors (can be ignored)

```{r}
mt_plot(mtdata, use = "tn_trajectories", use2 = "data", color = "Trial")
```


#### CALCULATE DERIVATIVES, MEASURES & DEVIATIONS ####
# A series of calculations made on the trajectories to provide useful data that are added to the
# mouse tracking object

```{r}
mtdata = mt_derivatives(
  mtdata,
  use = "tn_trajectories",
  save_as = "der_trajectories",
  dimensions = c("xpos", "ypos"),
  timestamps = "timestamps",
  prefix = "",
  verbose = FALSE
)
```


```{r}
mtdata <- mt_measures(
  mtdata,
  use = "der_trajectories",
  save_as = "measures",
  flip_threshold = 0,
  verbose = TRUE
)
```

```{r}
mtdata <- mt_deviations(
  mtdata,
  use = "tn_trajectories")  # RT : Response Time in "time stamp" units (i.e. frames = ~1/60th of seconds)
  # MD_above : Maximum Deviation from the straight path between start and end position
  # AUC : Area Under the Curve (between the trajectory and the straight line from start to end position)
d = data.frame(d,mt_export_wide(
  mtdata,
  use = "measures",
  use_variables = c("RT","MD_above","AUC")))
```



  
# STATISTICAL ANALYSIS
The data for this experiment are yet to be collected. But the statistical analysis should go as follows:
A series of linear mixed effect models will be run with each of the mouse tracking measures as dependent variable,
experimental condition as predictor, and participant and item as random factors

## Load the package for building Linear Mixed Effect Models object in R

```{r}
library("lme4")
```

## Build the models

```{r}
mRT = lmer(RT~Condition + (1|Participant) + (1|Item), data = d)
mMD = lmer(MD_above~Condition + (1|Participant) + (1|Item), data = d)
mAUC = lmer(AUC~Condition + (1|Participant) + (1|Item), data = d)
```



## Observe the models

```{r}
summary(mRT)
summary(mMD)
summary(mAUC)
```



## Other predictors may be added to the statistical analysis by building more complex models. For example, the model below adds another hypothetical fixed factor ("Predictor2") to the mAUC model.
```{r}
mAUC2 = lmer(AUC~Condition + Predictor2 + (1|Participant) + (1|Item), data = d)
```


# A model comparison (log likelihood ratio test) can assess if "Predictor2" significantly improves the model's predictions
```{r}
anova(mAUC, mAUC2)
```



# If significant, a more complex model including the interaction between the two fixed factors can be created

```{r}
mAUC2 = lmer(AUC~Condition * Predictor2 + (1|Participant) + (1|Item), data = d)
```


# another model comparison can assess if the interaction significantly improves the model's predictions
```{r}
anova(mAUC2, mAUC3)
```



